---
title: "Practical Machine Learning-Final Assignment"
author: "Morteza"
date: "July 2, 2017"
output: html_document
---

## Business Problem
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

## Understanding Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project comes from http://groupware.les.inf.puc-rio.br/har. Full source:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. "Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13)". Stuttgart, Germany: ACM SIGCHI, 2013.

Data description (from the authors' website):

"Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg)."

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
set.seed(12345)
```

```{r}
Train_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Test_url  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(Train_url))
testing  <- read.csv(url(Test_url))

dim(training)
dim(testing)
```

## Preparing Data

```{r}
# remove variables with Zero variance and those variables with more than 95% NA data

NZV <- nearZeroVar(training)
training <- training[, -NZV]
dim(training)

Data_NA    <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, Data_NA==FALSE]
dim(training)


# remove identification only variables (columns 1 to 5)
training <- training[, -(1:7)]

# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
```
## Developing Predictive Model


```{r}
# model fit
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFit <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFit$finalModel
Varimportance <- varImp(modFit, scale=TRUE)
dotPlot(Varimportance, top = min(20, dim(Varimportance$importance)[1]))


# prediction on Test dataset
predict <- predict(modFit, newdata=TestSet)
confMat <- confusionMatrix(predict, TestSet$classe)
confMat

# plot matrix results
plot(confMat$table, col = confMat$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMat$overall['Accuracy'], 4)))
```

## Applying the selected model to the test data set

```{r}
predictTEST <- predict(modFit, newdata=testing)
predictTEST
```


